---
title: "SMN-risk: an update from the Hodgkin Lymphoma survivorship cohort"
author: "Gordon P. Watt, PhD"
date: "`Sys.Date()`"
execute:
  echo: false
  warning: false
  error: false
  message: false
format: docx


---

```{r}

#| label: load-packages

library(tidyverse)
library(ggplot2)
library(readxl)
library(stringr)
library(survival)
library(ggsurvfit)
library(tidycmprsk)
library(Hmisc)
library(gt)
library(ggpubr)
library(gtsummary)
library(flextable)
library(cmprsk)

knitr::opts_chunk$set(warning = F, error = F, message = F, echo = F)

```

```{r}

#| label: read-derived-data

# Derived data are created by scripts/01_clean-data.R; run that first
# whenever the raw export or the variable definitions change.

source("scripts/00_paths.R")

d_desc <- readRDS(derived_anly)

cr1 = cmprsk::cuminc(d_desc$ftime_cmpr_inv, d_desc$cmpr_inv)

bmi_miss = table(d_desc$bmi_fup == 99)[2]

```


# INTRODUCTION

The purpose of this analysis report is to describe the incidence of subsequent malignant neoplasms (SMNs) in a cohort of Hodgkin Lymphoma survivors treated in 1989 or later. 

# METHODS

Follow-up time began 5 years after diagnosis month/years. In the primary analysis, follow-up time was calculated as the time from start of follow-up until the diagnosis of an invasive SMN, death, or end of follow-up. Death was considered a competing risk in calculation of cumulative incidence and multivariable hazard models.

Secondary analysis: (1) for female participants, hazard of 1st invasive breast cancer with censoring at diagnosis of other invasive cancers; (2) for male participants, hazard of 1st invasive prostate cancer;... 

Notes: BMI missing for `r bmi_miss` participants; not considered for remaining analysis.


```{r}

#| label: Description of study population


# labs_desc = list(
#   age_dx_cat,
#   y_dx_cat,
#   sex_rec,
#   any_chemo,
#   anth_rec,
#   alk_rec,
#   ctype_combi,
#   any_rt,
#   rx_combi,
#   tot_fuptime_cat,
#   num_inv_smn,
#   num_smn
# )


desc_labels = list(
  age_dx_cat ~ "Age at HL diagnosis",
  y_dx_cat ~ "Year of HL diagnosis",
  tot_fuptime_cat ~ "Total follow-up time, years",
  sex_rec ~ "Sex",
  smk_evr_rec = "Ever smoked",
  ctype_combi ~ "Chemotherapy/type received",
  alk_rec ~ "Alkylating agents received",
  anthra_rec ~ "Anthracyclines received",
  any_rt ~ "Any RT received for HL",
  rx_combi ~ "Treatment combination for HL",
  num_inv_smn ~ "Total number of invasive SMNs",
  num_smn ~ "Total number of subsequent cancers"
)

t_desc = d_desc %>% select(
  age_dx_cat,
  y_dx_cat,
  tot_fuptime_cat, 
  sex_rec,
  smk_evr_rec,
  rx_combi,
  alk_rec,
  anthra_rec,
  ctype_combi,
  any_rt,
  num_inv_smn,
  num_smn
) %>% gtsummary::tbl_summary(
  type = c("any_rt", "smk_evr_rec") ~ "categorical",
  label= desc_labels
  ) %>% modify_footnote(
    everything() ~ NA, abbreviation = F
  ) %>% bold_labels()


```

**Table 1**, below, displays the characteristics of the study population. Diagnosis years ranged from 1989 through 2013, with 48% treated in 2000 or later. Diagnosis age ranged from 4 to 79, with median follow-up time of 12 years (range 0.1-28) conditional on survival to 5 years after diagnosis. Sixty-seven percent received both RT and chemotherapy for primary treatment; for those receiving chemotherapy, 70% received both anthracyclines and alkylators.Sixteen percent received at least one invasive SMN diagnosis; 3% had more than 1 invasive SMN. 

```{r}

#| label: render-desc-tbl

t_desc %>% as_flex_table(
) %>% add_header_lines(
  "Table 1. Descriptive statistics of Hodgkin Lymphoma survivors diagnosed 1989-2013"
) %>% add_footer_lines(
  "Abbreviations. HL, Hodgkin Lymphoma; SMN, subsequent malignant neoplasm; RT, radiation therapy"
) %>% footnote(
  part = "body",
  i = 12, j = 1,
  value = as_paragraph(c(
    "Follow-up time was calculated from 5 years after diagnosis until diagnosis of an invasive SMN, death, or end of follow-up")),
    ref_symbols = c("1")
  ) %>% bold(
    i = 1, part = "header"
  ) %>% flextable::autofit()




```

```{r}
#| label: rx-yr-crosstab

cross_labels = list(
  rx_combi ~ "Treatment combination",
  y_dx_cat ~ "Year of diagnosis",
  ctype_combi ~ "Chemotherapy receipt and type"
)

cross1 = d_desc %>% gtsummary::tbl_cross(
  row = rx_combi,
  col = y_dx_cat,
  percent = "column",
  label = list(
    rx_combi ~ "Treatment combination",
    y_dx_cat ~ "Year of diagnosis")
)

cross2 = d_desc %>% gtsummary::tbl_cross(
  row = ctype_combi,
  col = y_dx_cat,
  percent = "column",
  label = list(
  y_dx_cat ~ "Year of diagnosis",
  ctype_combi ~ "Chemotherapy classes"
)
)

out1 = tbl_stack(
  list(
    cross1, cross2
  )
)  %>% bold_labels()

```

**Table 2** displays the distribution of treatment combinations across years of diagnosis. RT was excluded from 20% of treatment plans overall, ranging from 11% in 1989-1994 to 30% in 2005-2013. Combined anthracycline+alkylator agent therapy was given in 55% of treatment plans from 1989-1994 and in 85% of plans in 2005-2013.

```{r}

#| label: render-cross

out1 %>% as_flex_table(
) %>% add_header_lines(
  "Table 2. Treatment types by year of diagnosis"
) %>% footnote(
  i = 10, j = 1,
  value = as_paragraph(
    "Chemotherapy received but not classified as anthracycline-based or alkylator-based"
  )
) %>% bold(
  part = "header"
) %>% flextable::autofit()

```

\newpage


```{r}


#| label: cuminc-1st-inv

# filter to invasive smnss

cuminc_1st_inv = tidycmprsk::cuminc(Surv(ftime_cmpr_inv, as.factor(cmpr_inv)) ~ 1, data = d_desc)

p1 = cuminc_1st_inv %>% ggcuminc(
) + add_confidence_interval(
) + add_risktable(
) + scale_ggsurvfit(
  y_scales = list(
    breaks = seq(0,1,by = .1),
    limits = c(0,1)
  )
) + labs(
  x = "Follow-up time, years",
  y = "Cumulative incidence",
  title = "Overall"
) + theme(
  legend.position = "right"
)

cuminc_1st_inv_ydx = tidycmprsk::cuminc(Surv(ftime_cmpr_inv, as.factor(cmpr_inv)) ~ y_dx_cat, data = d_desc)

# manual modification removing outliers (due to calculation of pt with months/years)
cuminc_1st_inv_ydx$tidy = cuminc_1st_inv_ydx$tidy[cuminc_1st_inv_ydx$tidy$estimate < 0.5,]

p2 = cuminc_1st_inv_ydx %>% ggcuminc(
) + add_confidence_interval(
) + add_risktable(
) + scale_ggsurvfit(
  y_scales = list(
    breaks = seq(0,1,by = .1),
    limits = c(0,1)
  )
) + labs(
  x = "Follow-up time, years",
  y = "Cumulative incidence",
  title = "By diagnosis year"
) + theme(
  legend.position = "right"
)


cuminc_est = tbl_cuminc(cuminc_1st_inv)

```

*Figure 1* displays the cumulative incidence curves for any invasive SMN accounting for death as a competing risk. The cumulative incidence of any invasive SMN was `{r} cuminc_est$table_body$stat_1` at 10 years and `{r} cuminc_est$table_body$stat_2` at 20 years (Figure 1A). There was no statistical evidence of changes in SMN incidence across diagnosis years (p=0.28).

```{r, fig.height=8}


ggpubr::ggarrange(p1,p2,
                  nrow = 2,
                  common.legend = T,
                  labels = c("A","B"),
                  legend = "bottom"                 )

```

